Fractional Types

### 2019-11-15 fracGC/p.tex
working on sec 4

A conventional way to ensure that every allocation is paired with a
matching de-allocation is to impose a stack discipline. An
illustrative demonstration of this idea is the \verb|with_ancilla|
operator from Quipper~\cite{Green:2013:QSQ:2491956.2462177}:
\begin{verbatim}
with_ancilla :: (Qubit -> Circ a) -> Circ a
\end{verbatim}
The operator takes a block of gates parameterized by an ancilla,
allocates a new ancilla initialized to $\ket{0}$, and runs the given
block of gates. At the end of its execution, the block is expected to
return the ancilla to the state $\ket{0}$ at which point it is
de-allocated. In this case, allocation and de-allocation of ancillae
requires nothing beyond conventional parameter-passing techniques in
which the parameter is allocated before entry to the function and
de-allocated at exit.

This scoped model is therefore quite a pragmatic choice. It is however
limited and to understand its limitations more vividly, we propose the
following analogy with money.

@@

>>>>

A conventional and efficient way to ensure that every allocation is
paired with a matching de-allocation is to impose a stack
discipline. An illustrative demonstration of this idea is the
\verb|with_ancilla| operator from
Quipper~\cite{Green:2013:QSQ:2491956.2462177}:
\begin{verbatim}
with_ancilla :: (Qubit -> Circ a) -> Circ a
\end{verbatim}
The operator takes a block of gates parameterized by an ancilla,
allocates a new ancilla initialized to $\ket{0}$, and runs the given
block of gates. At the end of its execution, the block is expected to
return the ancilla to the state $\ket{0}$ (this is the second
subproblem addressed later) at which point it is de-allocated. In this
scoped-by-construction approach, allocation and de-allocation of
ancillae requires nothing beyond conventional parameter-passing
techniques in which the parameter is allocated before entry to a
function and de-allocated at exit.

This scoped model is therefore quite a pragmatic choice. It is however
limited and to understand its limitations more vividly, we propose the
following analogy: allocating an ancilla by creating a new wire in the
circuit is like borrowing some money; the computation has access to a
new resource temporarily. De-allocating the ancilla is like returning
the borrowed money; the computation no longer has access to the
additional resource. It would be unreasonably restrictive to insist
that if the person (function) borrowing the money must be the same
person (function) who returns it, i.e., preventing the following
computation:
\begin{verbatim}
  |-----\  /-----|
         \/
        / \
  |----/   \-----|
\end{verbatim}
Indeed, as far as reversible or quantum computation is concerned, the
circuit above is perfectly sensible since it conserves information
(all money borrowed is repaid and the net balance is 0). In fact one
can go further, if one person borrows \$10, why couldn't three
different persons collectively collaborate to pay back the debt with
one person paying \$5, another \$2, and a third \$3. This extra
generality is not a gratuitous concern: since scope is fundamentally a
static property of programs, it does not allow the flexibility of heap
allocation in which the lifetime of resources is dynamically
determined. To summarize, since the reversible-quantum computational
framework guarantees that information is preserved, it also permits
fascinating mechanisms in which computations can be sliced and
diced, decomposed and recomposed, run forwards and backwards, in
arbitrary ways as long as the net balance is 0.

@@

%%%%% 
\subsection{Pointed Types}
